---
title: "LDA assignment"
author: "Ted Laderas (and you!)"
date: "May 23, 2016"
output: html_document
---

In this assignment, you will run the Linear Discriminant Analysis (LDA) algorithm on the Wisconsin Breast Cancer Dataset. You will use this Markdown document to record your observations and code.

Take a look at the features that are provided in the data (there is a Readme in the `/data` folder). Your first step will be to decide what features you will use and justify it based on your knowledge about the problem. If you'd like more info, you can read the paper about how the data was generated here: [Nuclear Feature Extraction For Breast Cancer Tumor Diagnosis](https://www.researchgate.net/publication/2512520_Nuclear_Feature_Extraction_For_Breast_Tumor_Diagnosis). What features do you think will distinguish best between Malignant and Benign cancer cells?

The first step is to load the data (in the `/data/` folder). What kind of data is it? What command will you use? Don't forget to load any libraries you might need. If you need help or get stuck, please note that there is a Hint Sheet called `hintSheet.Rmd` in the folder. 

```{r}
##Write your data loading code here
##be sure to read the readme file as well. 
##Do some EDA as well and show any relevant plots

##Some questions: What are the classes? What is the distribution of classes in the training set and the test set?
##Are there any concerns?
```

Based on your reading of the dataset, you may want to select some features to include and some to exclude. (Hint: you can use the `select()` command from the `dplyr` package to select particular features.)

```{r}

```

Now you can train your learner on the training data. What is the class that we're training on? How do we do this? Adapt the code that we used on the Iris dataset here. 

```{r}
##Write your lda code here.
```

Look at any LDA diagnostic plots. Are there any issues with the LDA analysis? How good is the discriminant?

```{r}
##write any plotting code here.
```

Evaluate your learner on the test set. How well did it do? Does the test data give you any confidence that you built a good learner? What is your interpretation of the results? Do you think the problem is an easy or hard one? How did the learner do?

```{r}
##write any evaluation code here.
```

If you have time. Compare your results on classifying the test set with both algorithms (LDA and Classification Trees). Is there a way to see compare the misclassifcations? Did both algorithms do equally well?